---
title: "assignment_06"
author: "Stephen Smitshoek"
date: "05/05/2022"
output: pdf_document
---

```{r echo=FALSE}
setwd("C:\\Users\\sksmi\\PeytoAccess\\Personal\\Bellevue\\DSC520\\dsc520")
heights_df <- read.csv("data\\r4ds\\heights.csv")
library(ggplot2)
```

```{r echo=FALSE}
age_lm <-  lm(earn ~ age, data=heights_df)
summary(age_lm)
```
```{r echo=FALSE}
age_predict_df <- data.frame(earn = predict(age_lm, heights_df[,"age",drop=FALSE]), age=heights_df$age)

ggplot(data = age_lm, aes(y = earn, x = age)) +
  geom_point(color='blue') +
  geom_line(color='red',data = age_predict_df, aes(y=earn, x=age))
```

Mean Earn:
```{r echo=FALSE}
mean_earn <- mean(heights_df$earn)
mean_earn
```

Corrected Sum of Squares Total:
```{r echo=FALSE}
sst <- sum((mean_earn - heights_df$earn)^2)
sst
```

Corrected Sum of Squares for Model:
```{r echo=FALSE}
ssm <- sum((mean_earn - age_predict_df$earn)^2)
ssm
```

Residuals:
```{r echo=FALSE}
residuals <- heights_df$earn - age_predict_df$earn
residuals
```

Sum of Squares for Error:
```{r echo=FALSE}
sse <- sum(residuals^2)
sse
```

R Squared R^2:
```{r echo=FALSE}
r_squared <- ssm/sst
r_squared
```

Number of observations:
```{r echo=FALSE}
n <- length(heights_df$age)
n
```

Number of regression parameters:
```{r echo=FALSE}
p <- 2
p
```

Corrected Degrees of Freedom for Model:
```{r echo=FALSE}
dfm <- p - 1
dfm
```

Degrees of Freedom for Error:
```{r echo=FALSE}
dfe <- n - p
dfe
```

Corrected Degrees of Freedom Total:
```{r echo=FALSE}
dft <- n - 1
dft
```

Mean of Squares for Model:
```{r echo=FALSE}
msm <- ssm / dfm
msm
```

Mean of Squares for Error:
```{r echo=FALSE}
mse <- sse / dfe
mse
```

Mean of Squares Total:
```{r echo=FALSE}
mst <- sst / dft
mst
```

F Statistic:
```{r echo=FALSE}
f_score <- msm / mse
f_score
```

Adjusted R Squared R2:
```{r echo=FALSE}
adjusted_r_squared <- 1 - (1 - r_squared) * (n - 1) / (n - p)
adjusted_r_squared
```

p-value:
```{r echo=FALSE}
p_value <- pf(f_score, dfm, dft, lower.tail=F)
p_value
```