---
title: "assignment_07"
author: "Stephen Smitshoek"
date: "05/05/2022"
output: pdf_document
---

```{r echo=FALSE}
setwd("C:\\Users\\sksmi\\PeytoAccess\\Personal\\Bellevue\\DSC520\\dsc520")
heights_df <- read.csv("data/r4ds/heights.csv")
```

```{r echo=FALSE}
earn_lm <-  lm(earn ~ height + sex + ed + age + race, data=heights_df)
summary(earn_lm)
```

```{r echo=FALSE}
predicted_df <- data.frame(
  earn = predict(earn_lm, heights_df),
  ed=heights_df$ed, race=heights_df$race, height=heights_df$height,
  age=heights_df$age, sex=heights_df$sex
  )
```

Mean Earn:
```{r echo=FALSE}
mean_earn <- mean(heights_df$earn)
mean_earn
```

Corrected Sum of Squares Total:
```{r echo=FALSE}
sst <- sum((mean_earn - heights_df$earn)^2)
sst
```

Corrected Sum of Squares for Model:
```{r echo=FALSE}
ssm <- sum((mean_earn - predicted_df$earn)^2)
ssm
```

Residuals:
```{r echo=FALSE}
residuals <- heights_df$earn - predicted_df$earn
residuals
```

Sum of Squares for Error:
```{r echo=FALSE}
sse <- sum(residuals^2)
sse
```

R Squared:
```{r echo=FALSE}
r_squared <- ssm/sst
r_squared
```

Number of observations:
```{r echo=FALSE}
n <- NROW(heights_df)
n
```

Number of regression parameters:
```{r echo=FALSE}
p <- 8
p
```

Corrected Degrees of Freedom for Model:
```{r echo=FALSE}
dfm <- p - 1
dfm
```

Degrees of Freedom for Error:
```{r echo=FALSE}
dfe <- n - p
dfe
```

Corrected Degrees of Freedom Total:
```{r echo=FALSE}
dft <- n - 1
dft
```

Mean of Squares for Model:
```{r echo=FALSE}
msm <- ssm / dfm
msm
```

Mean of Squares for Error:
```{r echo=FALSE}
mse <- sse / dfe
mse
```

Mean of Squares Total:
```{r echo=FALSE}
mst <- sst / dft
mst
```

F Statistic
```{r echo=FALSE}
f_score <- msm / mse
f_score
```

Adjusted R Squared:
```{r echo=FALSE}
adjusted_r_squared <- 1 - (1 - r_squared) * (n - 1) / (n - p)
adjusted_r_squared
```