---
title: "Play Time"
author: "Jared P. Lander"
date: "18/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Title - Also a Heading 1

_this will be italicized_

__this will be bolded__

## Heading 2

Build an unordered list

- Item 1
- Item 2
- Item 3


Build an ordered list mixed with an unordered list

1. An item
1. Another item
    - Sublist item
    - Another sublist item
    - one more item
1. Another ordered item

The following is a link
[My Website](http://www.jaradlander.com)

## Another Heading 2
This inserts an image

#### Heading 4
A regular equation
$$
  \boldsymbol{\hat{\beta}} = (X^TX)^{-1}X^TY
$$

An inline equation: $\bar{x}=\frac{1}{n}\nsum_(i=1)^n$ with no spaces

### Heading 3
> This is the start of a block quote
>
> This is the following line in the block quote

```{r code-and-plot}

#this is a comment
1+1
plot(1:10)
2+2
```

```{r kable}
knitr::kable(head(iris), caption='Tabular data printed using kable.')
```

```{r DT}
library(DT)
data(diamonds, package='ggplot2')
datatable(head(diamonds, 100), 
          rownames=FALSE, extensions='Scroller', filter='top',
          options = list(
            dom = "tiS", scrollX=TRUE,
            scrollY = 400,
            scrollCollapse = TRUE
          )) %>%
  formatCurrency('price', digits=0) %>%
  formatStyle(columns='cut', valueColumns='cut', target='row',
               backgroundColor=styleEqual(levels=c('Good', 'Ideal'),
                                          values=c('red', 'green')
                                          )
               )
```

```{r pizza-map}
library(jsonlite)
pizza <- fromJSON('http://www.jaredlander.com/data/PizzaFavorites.json')
pizza
```

```{r interactive-graph}
library(WDI)
library(dplyr)
gdp <- WDI(country=c("US", "CA", "SG", "IL"),
                    indicator=c("NY.GDP.PCAP.CD"),
                    start=1970, end=2011)
names(gdp) <- c("iso2c", "Country", "PerCapGDP", "Year")
head(gdp)
gdpWide <- gdp %>%
  dplyr::select(Country, Year, PerCapGDP) %>%
  tidyr::spread(key=Country, value=PerCapGDP)
head(gdpWide)
library(dygraphs)
dygraph(gdpWide, main='Yearly Per Capita GDP',
        xlab='Year', ylab='Per Capita GDP') %>%
  dyOptions(drawPoints = TRUE, pointSize = 1) %>%
  dyLegend(width=400) %>%
  dyRangeSelector(dateWindow = c("1990", "2000"))
```
